### 2020-07-16 paper.Rmd
added footnote

A motivating example illustrates that aiming to simultaneously optimize 
similarity of means and variances at the same time is not only 
desirable, but may also be feasible. I created 14 data points from a 
univariate normal distribution ($M = 15$, $SD = 3$) and generated all `r 
length(partitions)` possible ways to partition the 14 data points into 
$K = 2$ equal-sized groups. I computed the difference in the two means 
and two variances for each partition. This simple difference metric was 
chosen for purposes of interpretability; in fact, there is a a 1-to-1 
correspondence between these differences and $SSE$ and 
$SSE_{\mathrm{VAR}}$, respectively (Figure 1). Figure 2 illustrates 
similarity in means plotted against similarity in variances across all 
partitions. Ideally, a partition is chosen that minimizes the 
difference with regard to means and variances (see the triangular 
symbol). However, optimal k-means anticlustering would create two groups 
that are not particularly similar with regard to their variance and 
chose the partition that is represented by the squared symbol. The 
diamond symbol represents that partition that where the two groups have 
the least difference in variances. The black symbols illustrate the 
other members of the optimal pareto set.

>>>>

A motivating example illustrates that aiming to simultaneously optimize 
similarity of means and variances at the same time is not only 
desirable, but may also be feasible. I created 14 data points from a 
univariate normal distribution ($M = 15$, $SD = 3$) and generated all `r 
length(partitions)` possible ways to partition the 14 data points into 
$K = 2$ equal-sized groups. I computed the difference in the two means 
and two variances for each partition. This simple difference metric was 
chosen for purposes of interpretability; in fact, there is a a 1-to-1 
correspondence between these differences and $SSE$ and 
$SSE_{\mathrm{VAR}}$, respectively (Figure 1).[^onlyonegroup] Figure 2 
illustrates similarity in means plotted against similarity in variances 
across all partitions. Ideally, a partition is chosen that minimizes 
the difference with regard to means and variances (see the triangular 
symbol). However, optimal k-means anticlustering would create two groups 
that are not particularly similar with regard to their variance and 
chose the partition that is represented by the squared symbol. The 
diamond symbol represents that partition that where the two groups have 
the least difference in variances. The black symbols illustrate the 
other members of the optimal pareto set.

[^onlyonegroup]: Note that this such a correspondence no longer holds 
when there are more than two groups or there is more than one 
numeric attribute (i.e., if $K > 2$ or $P > 1$).
